### 2016-06-16 exam2.tex
more fleshed out outline + more intro folded into latex

\documentclass[letterpaper,onecolumn,titlepage]{Ythesis}
\usepackage{siunitx}
\usepackage{amsmath}
\usepackage{url}
\sisetup{detect-weight=true, detect-family=true}
\newcommand{\micron}[1] {\SI{#1}{\micro\meter}}

%\usepackage{setspace}
%\onehalfspace

\title{Visualizing Relationships Amongst and With Climate Ensembles}
\author{Hannah Aizenman}
\committee{ Dr. Michael Grossberg(Advisor), Robert Haralick, Huy Vo}
\submitted{}
\abstract{Climate datasets often inherently try to encapsulate complex systems. The variables are often high dimensional and often have long ranging interactions both temporally and spatially. Ensemble climate models often add another layer to the complexity because they add uncertainty to the already complex interactions. Scientists often try to visualize these interactions either by attempting to encode many if them in a visualization, using a machine learning algorithm to model the interaction and traditional approaches to visualize them, or building out a complex toolkit that relies on users to compare graphs of each of the sub interactions to build out the relationships. Tamara'Munzner's domain-task-idiom framework can be an effective way of grouping the visualization techniques that have become dominant in the ensemble evaluation community and understanding what it is they're trying to accomplish.}
\begin{document}
\makefrontmatter

\section{Introduction}

%Large datasets of climate and weather data are being produced from multiple sources. With cheaper and increased computer power, and cheaper disk storage, it is now possible to more easily create long model simulations at higher resolutions. Increasingly  high 
%resolution satellite data and reanalysis projects which blend both model prediction and measurements create readily available large uniformly sampled spatio-temporal resolution datasets. Measured data from networks of ground stations and proxy data from trees, 
%glaciers, oceans, corals, fossils, and historical records are also becoming more accessible, 
%yielding sometimes irregularly sampled data that provides a host of 
%information on climate variability. The great size of many of these data sets mean that just providing a download link does not really make the data accessible to the public or even other scientists. It is important for the scientists who produce the data to have a way to make it explorable..

And that those explorations are understandable...while many papers(\cite{Munzner} \cite{Acquired Codes} \cite{...}discuss how more complex tasks inherently encourage more complex visualizations, they also warn that more complex visual idioms are often more difficult to understand and so they may not be beneficial. Some authors though discuss how highly specialized dashboards may be more productive once specialists are trained on the software.

Organization:
\begin{table}
data & task & idiom  & AUDIENCE\\

\end{table}

\subsection{Scope}

This paper focuses on the visualization of:
\begin{itemize}
\item climate and weather data - observational and model
\item has both a spaitial and temporal component
\item requires aggregation(ML or stats) before can be visualized with simple viz (line, bar, pie, scatter, heatmap)
\end{itemize}

%%some sorta transition from the generic to the specific tools-and shift this to be more tool oriented

Munzner \cite{Munzner14} defines a time-varying dataset as one in which time is an intrinsic attribute of how the various observations are measured; for example recording snow fall every 3 hours or stock prices at the beginning or end of day. She contrasts this with datasets that are amassed over time where the length of the record does not intrinsically mean the data is time varying; for example [insert something that isn't horse racing]

Often the observations in these time-varying datasets are also spatially varying, spreading over the earth or the brain or another variable space where all the observations in the high dimensional space are of the same unit. %%note: must get some reference/notation/way to discuss this

This poses a difficult visualization task because the researcher is trying to capture multiple levels of interaction:
\begin{enumerate}
	\item intraobservational: snowfall in New York and snowfall in New Jersey on August 9th 2015)
	\item interobservational: global snowfall on August 9th and August 10th
	\item a mixture thereof: does snowfall in New York on the 9th affect snowfall in New Jersey on the 10th?
\end{enumerate}

\section{Simple idioms, hard math}

\subsection{Probabilistic aggregations/information gain metrics}
\label{sec:probabilistic}
%%Dump Nir into here-it kinda ties into the band depth stuff that'll show up later
\subsection{markov chain analysis}
\subsection{clustering}

\section{Let's throw Everything in one picture}                                                        
%% Evolution of Single Task/Single Idiom to Higher Order 
\subsection{Scatter, matrix of scatter, scagnostics}                                        
\subsection{Quiver, Flow, Spaghetti, Lasagna}
\subsection{BoxPlots, Functional Box Plots, Countour Box Plots}

Countour Boxplots \cite are a cross between boxplots and the  
\subsection{Maps & Heatmaps -> Glyphs}

\subsection{Let's Jam all the simple together as Toolkits}
Some research institutions provide a tool to explore data they created, such as the \cite{src:esrlpsd}, but since all these tools are site specific, a researcher would either have to use the datasets on the website or submit their datasets for uploading 
(assuming their dataset meets the submission criteria for an aggregation site). A scientist also has the option of using the CDAT suite of libraries, developed by the \cite{WilliamsEtAl13}, GrADS, \cite{src:grads},and Ferret, \cite{src:HankinEtAl96}. The major critique with all of these though is that they mostly render the data as is, with limited support for statistical analysis and the application of typical meterological analysis. Support for even common machine learning techniques often requires extending the library; therefore for the purpose of this discussion it is somewhat useful to treat these tools more as plotting libraries than visualization systems. %%somewhere scope out/differentiate between plotting and viz

\section{Conclusion}
\label{sec:conclusion}

\pagebreak
\bibliographystyle{abbrv}
\bibliography{exam2}

\end{document}