### 2016-09-29 sections/section0d.tex
new outline

\documentclass[../main.tex]{subfiles}

\begin{document}
\section{Independent Observations (0D fields)}

\subsection{Univariate Observations}
%%histogram and pdfs/distribution fitting

\subsection{Bivariate Observations}
%%introduce box plot

\subsection{Multivariate Observations}
%% large scale box: letter, spark

\subsection{Relationships Between Variables}
\subsubsection{Bivariate}
%%scatter segue into bagplot and HDR?

\subsection{Multivariate Observations}

%% radial, parallel coordinates, factor plotl
%%high d-PCA, ICA, markov