### 2016-09-29 sections/section1d.tex

\documentclass[../main.tex]{subfiles}

\begin{document}
\section{1D Fields}

\subsection{Independent Variables}

\subsection{Univariate Observations}
%%functional box + derivatives,

\subsection{Bivariate Observations}
%%hurricane tracks (x+y dept)

\subsection{Multivariate Observations}
LDA, CCA, 
\subsection{Relationships Between Variables}

\subsubsection{Bivariate}
surface map, functional correlations, planes?
\subsection{Multivariate Observations}
PLS, functional PCA