### 2016-09-29 sections/sectionnd.tex
new outline

\documentclass[../main.tex]{subfiles}

\begin{document}
\section{ND Fields}

\subsection{Independent Variables}
\subsection{Univariate Observations}
subspace decomposition, projection persuite, manifolds,
spation temporal viz, yield curve
\subsection{Bivariate Observations}

\subsection{Relationships Between Variables}
\subsubsection{Bivariate}
\subsection{Multivariate Observations}