### 2017-05-09 sections/sectionnd.tex
somehow lost my sections...

\documentclass[../main.tex]{subfiles}

\begin{document}
\section{ND Fields}
Toolkits, all the Toolkits.

overview-zoom/filter-details-on-demand is the guiding mantra for many of the
tools built to visiaulzie demnsionally complex data \cite{shneiderman1996eyes}
because these tools facilitate engagment with the data at the aggregate and
encourage drilling down to the core. Some of the most widley used of these
tools are:
*parview
*vistrails
*...

\subsection{Independent Variables}
\subsection{Univariate Observations}
Toolkits, all the Toolkits.

%cite my calender paper here!
subspace decomposition, projection persuite, manifolds,
spation temporal viz, yield curve
\subsection{Bivariate Observations}

\subsection{Relationships Between Variables}
\subsubsection{Bivariate}
\subsection{Multivariate Observations}