### 2017-06-23 paper/sections/discrete.tex
scraps of start of categorical

\documentclass[../main.text]{subfiles}
\begin{document}

%%% throw up a visualization of what this sort of data means

Demographic data is some of the most commonly analyzed in any number of fields,
and often researchers are interested in the joint distributions of different
demographic categories (such as race, income level, and education level). While
a table is encouraged for a small number of variables
\cite{munznertablesgood}), it can get unweildy rather quickly.

Almost regardless of the number of categories, a single categorical variable
is visualized as a bar graph of the frequency of each catgory \cite{friendly:...}

More frequencies? Fourfold display \cite{Fienberg, 1975, Friendly, 1994a,c},
and radial plots \cite{...,....}

Correlational maps....

\cite{friendly:visualizingcategoricaldata}

Categorical attributes can also be mapped to mathematical sets such that an
observation's categorical attributes can also be a set of sets and visualized
as the distribution of set intersections \cite{UpSet, parrallel sets paper}

\end{document}