### 2017-08-19 paper/sections/multivariate.tex
title + consolidation

\begin{figure}

\end{figure}
The parallel coordinates plot \cite{islenberg} shows how different quantative
values are related to each other conditioned on a categorical variable whose
categories are mapped to different colors. As shown in figure~\ref{fig: